A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-30 manuscript.Rmd

<!--AB: TODO add a quick summary -->

@@

PACs come with a challenge though.
In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->

@@

<!--AB: two lines of what the future of the package, of such standards, and prereigstration is. Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->

>>>>

## Summary

We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.

We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
These results are based on simulated data, so that there is no danger of peeking at the real data while writing the scripts.
This way, all researcher's degrees of freedom are eliminated.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration

## Limitations of repro

`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
We hope that despite the early development stage, `repro` can support researchers to work more reproducible.

The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
However, Git and Make are itself included in the Docker image created by `repro`.
Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
In other words the only hard requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
The Docker approach has two vulnerabilities.
First, and more important, the Docker image for the project has to remain available.
The Dockerfile (the plain text description to build a Docker image) on the other hand is insufficient because it relies on to many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
The Docker image can be saved to disk and treated like any other file that requires careful archivation.
However, if the existence of the Docker image is guaranteed, researchers still require software to run the image.
To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
This process is technically possible and would guarantee reproducibility for decades without any software dependency.
However, this process requires much technical knowledge and is currently not supported by `repro`.
In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.
The Makefile, for example, is written in a way that researchers can manually trace the dependencies and execute commands in the right order, in case they are, for whatever reason, unable to run Make.
Similarly, even when Docker does not work, the Dockerfile still serves as unambiguous documentation of how the original system was set up and may help future users to create a software environment that closely resembles the original.

@@

We realize that PACs are challenging.
Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a complement.
The rigidness of PAC is only than to much of a good thing, if one holds the view that researchers can for no reason deviate from their preregistration.
However, we are convinced that many reasons justify changing the preregistered code, but are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.
PAC enables just that.
Researcher can change their plans and possibly defend their decision, e.g. in the discussion, while readers of the manuscript can clearly see what was changed through version control.

Another challenge is that PAC requires more programming expertise than a usual analysis.
In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->

@@

Open science practices are a continually evolving field where technical innovations foster changes in research practice.
Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.

This ideal of a fully automatic reproduction of research projects meets in practice a wide range of demands on how user-friendly and how powerful the software solutions are.
Some may find Make too complicated or that Docker requires too much storage.
Yet others may find that they require other programming languages or scale their computation across hundreds of computers, e.g. via high-performance computing clusters or cloud computing.

`repro` was designed modularly to meet such demands.
At the moment, `repro` only supports the combination of RMarkdown, Git, Make, and Docker.
However, there are alternatives for each of these elements that may fit better into an individual research project.
RMarkdown could be complemented or replaced by a dynamic Microsoft Word document with the help of `officer`[@officer] or `officedown`[@officedown] to accommodate a wider arrange of journal submission standards.
Instead of using formal version control with Git, `repro` could automatically save snapshots for increasing user-friendliness.
Make could be replaced by the more R-centered alternative `targets` for more convenience.
Docker could be combined with `renv` [@R-renv] to control the package versions precisely (our approach fixes the date, `renv` the exact package version).
Alternatively, Docker could be replaced by the more lightweight `renv` if no dependencies outside of R are considered crucial.
Docker does not satisfy the requirements of many HPC environments, but Singularity was designed to avoid this limitation while still being compatible with Docker images.

`repro`'s modular structure allows such alternative workflow, though they have not been implemented.
Depending on the demand by users, we will implement some of these alternative workflows in `repro`.
